---
title: "R Notebook"
output: html_notebook
---

CLEANED VERSION


```{r}

library(readr)
library(tidyverse)
library(stringr)
data2014 <- read_csv("data2014.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Abs = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2015 <- read_csv("data2015.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Abs = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2016 <- read_csv("data2016.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2017 <- read_csv("data2017.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2018 <- read_csv("data2018.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2019 <- read_csv("data2019.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))


data <- bind_rows(data2014,data2015, data2016, data2017, data2018, data2019)

Fourth_Plus <- c("4","OT1","OT2","OT3","OT4")

data <- data %>%
  filter(Quarter %in% Fourth_Plus) 

```



Add variable showing logical if play was home or away
```{r}
data$Is_Home <- if_else(data$Play == data$Home_Court,"Home", "Away")
```



```{r}
data$Play <- ''

#combine away and home play details into one column
for(i in 1:nrow(data)) { 
  if (is.na(data$Away[i])) {
    data$Play[i] <- data$Home[i]
  } else { 
    data$Play[i] <- data$Away[i]
  } 
}

#log shot results per play
data$Shot_Result <- NA
data[which(str_detect(data$Play, ".*misses.*")), "Shot_Result"] <- 0
data[which(str_detect(data$Play, ".*makes.*")), "Shot_Result"] <- 1


#Extract Player

data$Player <- str_extract(data$Play, "\\b\\p{Lu}\\.\\s*\\w+")
data$Shot_Length <- str_extract(data$Play, "..\\sft.*")
data$Shot_Length <- str_extract(data$Shot_Length, ".\\d")


#is dunk? might help to illinate from clutch SHOOTING as opposed to clutch playing

data$Is_Dunk <- str_extract(data$Play, " 2-pt dunk")
data$Is_Dunk <- if_else(is.na(data$Is_Dunk), "No", "Yes")

#is three?
data$Is_Three <- str_extract(data$Play, " 3-pt jump shot")
data$Is_Three <- if_else(is.na(data$Is_Three), "No", "Yes")

#is free throw
data$Is_FT <- str_extract(data$Play, " free throw")
data$Is_FT <- if_else(is.na(data$Is_FT), "No", "Yes")

#score differential
data$Away_Score <- as.numeric(data$Away_Score)
data$Home_Score <- as.numeric(data$Home_Score)

for(i in 1:nrow(data)) { 
  if (is.na(data$Away_Points_Added[i])) {
    data$Points_Added[i] <- data$Home_Points_Added[i]
  } else { 
    data$Points_Added[i] <- data$Away_Points_Added[i]
  } 
}

#margin at time of shot
for(i in 1:nrow(data)) { 
  if (data$Is_Home[i] == "Home") {
    data$Lead[i] <- data$Home_Score[i] - data$Away_Score[i]
  } else { 
    data$Lead[i] <- data$Away_Score[i] - data$Home_Score[i]
  } 
}

data$Points_Added[which(is.na(data$Points_Added))] <- 0
```



OK now to create the clutch analysis main data base. 4th quarter lead or deficit of less than three shots exluding FT


```{r}

clutch_analysis <- data %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4, Quarter %in% Fourth_Plus) %>%
  filter(Is_FT == "No") %>%
  select(Year, Round, Game, Player, Home_Score, Away_Score, Diff_Abs, Lead, Quarter, Minutes,Seconds,Shot_Result,Points_Added,Shot_Length,Is_Dunk,Is_Three,Is_FT, Is_Home, Home_Court, On_Road, Play)

clutch_analysis$Shot_Length[which(is.na(clutch_analysis$Shot_Length),)] <- 0
       
```

Build star categories

```{r}
stars <- c("S. Curry", "D. Lillard", "L. James","J. Harden", "K. Durant","K. Leonard", "R. Westbrook", "D. Wade","D. Nowitzki","D. Rose","T. Parker","T. Duncan", "K. Bryant", "A. Davis", "D. DeRozan","D. Mitchell","G. Antetokounmpo", "J. Butler", "J. Embid", "K. Garnett", "K. Irving", "K. Lowry","K. Thompson","K. Towns","K. Walker","N. Jokic","P. George","P. Pierce","V. Oladipo", "M. Ginobli", "B. Griffin", "B. Simmons", "C. McCollum")

stars_current <- c("S. Curry", "D. Lillard", "L. James","J. Harden", "K. Durant","K. Leonard","R. Westbrook") 

stars_yesterday <- c("D. Wade","D. Nowitzki","D. Rose","T. Parker","T. Duncan", "K. Bryant")

```

Who today, takes the most shots in clutch? (last 6 years)
```{r}
x <- clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 30) %>%
  group_by(Player) %>%
  summarise(sum =n())%>%
  arrange(desc(Player))


```



create diff databases for all times left

```{r}
cl_30 <- clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 30)

cl_1 <- clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 60)

cl_2 <- clutch_analysis %>%
  filter(Minutes < 2)

cl_3 <- clutch_analysis %>%
  filter(Minutes < 3)

cl_4 <- clutch_analysis %>%
  filter(Minutes < 4)

cl_5 <- clutch_analysis %>%
  filter(Minutes < 5)
```

Figure out shots attempted and made by bums per time

```{r}
# 30 seconds
# 1 minute
# 2 minutes
# 3 minutes
# 4 minutes
# 5 minutes

cl_30 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_30 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_1 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_1 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_2 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_2 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_3 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_3 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_4 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_4 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_5 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_5 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


```


Results for bums:
30: 67/156
1: 107/237
2: 163/394
3: 232/538
4: 281/703
5: 327/852


same thing for stars now to make sure they worse:  THIS IS ONLY FOR LAST 6 YEARS. NEED CAREER TOO

```{r}
cl_30 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_30 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))



cl_1 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_1 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))



cl_2 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_2 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))



cl_3 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_3 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))



cl_4 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_4 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))




cl_5 %>%
  filter(Player %in% stars) %>%
  summarize(n())

cl_5 %>%
  filter(Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
```


I used this info to insert stars and bums for each minute marker in the excel "current_past_clutch" so it's ready now for ggplots

```{r message=FALSE}
library(readxl)
library(directlabels)
Clutch_Complete <- read_excel("Clutch Complete.xlsx")

player_list <- Clutch_Complete %>%
  select(Player, Min, FGM,FGA) %>%
  filter(!is.na(FGA)) %>%
  group_by(Player, Min) %>%
  summarize(tot_made = sum(FGM), tot_att = sum(FGA)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

#initial graph hiding the identity of the bums
ggplot(player_list[player_list$Player %in% stars_current | player_list$Player == "Bums",], aes(Min, Percentage, color = Player))+
  scale_x_continuous(limits = c(-.4,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink1","black", "red2", "gold2", "brown4", "purple4", "blue", "yellow4"))+
  geom_dl(aes(label = Player), method = list(dl.combine("first.points"), cex = 0.8))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

#Reveal the almighty bums
g <- ggplot(player_list[player_list$Player %in% stars_current | player_list$Player == "Bums",], aes(Min, Percentage, color = Player))+
  scale_x_continuous(limits = c(-.40,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink2","black", "red2", "gold2", "firebrick4", "purple4", "blue", "yellow4"))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

direct.label(g, "first.qp")

p <- ggplot(player_list[!player_list$Player %in% stars_current,], aes(Min, Percentage, color = Player))+
  geom_smooth(method = "loess")+
    scale_x_continuous(limits = c(0,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink1","blue", "red2", "black", "purple4", "gray", "gray3"))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

direct.label(p, "first.qp")

```

show low sample sizes for 30 seconds:

```{r}
player_list %>%
  filter(Min == .5) %>%
  group_by(Player) %>%
  summarize(total = sum(tot_att))%>%
  arrange(desc(total))
  
```


After explaining low sample sizes for those above bums, combine them all:

```{r}

player_list$Type <- if_else(player_list$Player %in% stars,"Stars", "Bums")

player_list %>%
  filter(Min == .5) %>%
  group_by(Type) %>%
  mutate(Percent = sum(tot_made) / sum(tot_att)) %>%
  summarise(Avr = mean(Percent))

#USE IN ARTICLE

bum_star_graph <- player_list %>%
  filter(Player != "Stars") %>%
  group_by(Type, Min) %>%
  summarize(tot_made = sum(tot_made), tot_att = sum(tot_att)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

ggplot(bum_star_graph, aes(Min, Percentage, color = Type))+
  geom_smooth(se = FALSE, size = 2)+
  scale_y_continuous(limits = c(0,1), breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  scale_x_continuous(breaks = c(.5,1,2,3,4,5))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))
```

Same things as above only this time let's only focus on the BEST closers:

```{r}
super_stars <- c("S. Curry", "L. James", "K. Durant","K. Leonard", "D. Nowitzki","T. Duncan", "K. Bryant")

bum_star_graph <- player_list %>%
  filter(Player != "super_stars") %>%
  group_by(Type, Min) %>%
  summarize(tot_made = sum(tot_made), tot_att = sum(tot_att)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

ggplot(bum_star_graph, aes(Min, Percentage, color = Type))+
  geom_smooth(se = FALSE, size = 2)+
  scale_y_continuous(limits = c(0,1), breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  scale_x_continuous(breaks = c(.5,1,2,3,4,5))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Super 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))
```



Great plot. clearly shows that as minutes wane, clutch shooters (ACROSS THEIR CAREER) choke more than bums (IN LAST 6 YEARS) but with 5 min stars are better as expected. 

Now go into last minute - show each and every second. 

This plot shows that even in the last minute, bums are better in most cases. 

```{r}
clutch_analysis$Star_or_Bum <- NA
clutch_analysis$Star_or_Bum <- if_else(clutch_analysis$Player %in% stars,"Stars", "Bums")


last_min <- clutch_analysis %>%
  filter(Minutes == 0) %>%
  group_by(Star_or_Bum, Seconds) %>%
  summarize(tot_made = sum(Shot_Result), tot_att = n()) %>%
  mutate(Percentage = round(tot_made/tot_att,2))

ggplot(last_min, aes(Seconds, Percentage, fill = Star_or_Bum, color = Star_or_Bum))+
  geom_smooth(method = "loess", se = FALSE, size = 2)+
  scale_x_continuous(breaks = c(0,1,5,30,60))+
  scale_y_continuous(limits = c(0,1))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Final Minute")+
  ylab("Shooting Percentage")+
  xlab("Seconds Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

```



Dirk 

```{r}
Clutch_Complete %>%
  filter(!is.na(FGA)) %>%
  filter(Player == "D. Nowitzki") %>%
  group_by(Min) %>%
  summarise(tot.att = sum(FGA), tot.made = sum(FGM), tot.ave = round(tot.made / tot.att,2))
```



Re-evaluate where to mention this graph in article

```{r}
time_perc_stars <- clutch_analysis %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4) %>%
  filter(Quarter == "4" | 
           Quarter == "OT1" | Quarter == "OT2" | 
           Quarter == "OT3" | Quarter == "OT4") %>%
  filter(Is_FT == "No" & Player %in% stars) %>%
  group_by(Minutes) %>%
  summarise(n = n(), Percentage = mean(Shot_Result))

time_perc_bums <- clutch_analysis %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4) %>%
  filter(Quarter == "4" | 
           Quarter == "OT1" | Quarter == "OT2" | 
           Quarter == "OT3" | Quarter == "OT4") %>%
  filter(Is_FT == "No" & !Player %in% stars) %>%
  group_by(Minutes) %>%
  summarise(n = n(), Percentage = mean(Shot_Result))

ggplot(time_perc_stars, aes(Minutes, Percentage))+
  geom_bar(stat = "identity", fill = "blue")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))

ggplot(time_perc_bums, aes(Minutes, Percentage))+
  geom_bar(stat = "identity", fill = "deeppink2")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))

#BEATUFUL PLOT!!! JUST MESS WITH COLORS TO FIND BEST MATCH. 
ggplot(NULL, aes(Minutes, Percentage))+
  geom_bar(aes(fill = "Percentage"), stat = "identity", data = time_perc_stars, alpha = .4, fill = "springgreen4")+
  geom_bar(aes(fill = "Percentage"), stat = "identity", data = time_perc_bums, alpha = .4, fill = "deeppink2")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Final 12 Minutes")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

```

Despite a dip in the 7 minute and 9 minute range, stars outshot bums until there were less than 3 minutes to go overall. This reflects the earlier line chart with the overlap around the 3 mark. (this current graph is diff in that it is only looking at specific minutes, so the 3 minute bar doesn't include shots from 2 mins or less in calculation)


#Why do bums do better?

Since the crossover occurs under 3 minutes in %, use this as the parameter. All 14 stars used in this.
```{r}
#plot shows bums have so many more close range shots USE FOR ARTICLE 
clutch_analysis$Shot_Length <- as.numeric(clutch_analysis$Shot_Length)


```


```{r}
clutch_analysis %>%
  filter(Minutes < 3) %>%
  ggplot(aes(Shot_Length, fill = factor(Star_or_Bum)))+
  geom_bar(position = "dodge", alpha = .7)+
  scale_fill_manual("Stars Vs. Bums", values = c("Bums" = "deeppink2", "Stars" = "springgreen4"))+
  theme_classic()+
  ggtitle("Amount of Shots Per Foot \n Pooled 'Closers' vs. Pooled Bums in Playoffs When Tied or Down 3 or Less \n Final 3 Minutes")+
  scale_x_continuous(breaks = seq(0,62,2))+
  ylab("Number of Shots")+
  xlab("Length of Shots (in feet)")+
  theme(plot.title = element_text(hjust = 0.5))
```


