---
title: 'Analyzing The Most Clutch NBA Player Today: Insert Bum Here'
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    fig_height: 4.5
    fig_width: 7
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
---

# Loading Data

```{r}

library(readr)
library(tidyverse)
library(stringr)
data2014 <- read_csv("data2014.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Abs = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2015 <- read_csv("data2015.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Abs = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2016 <- read_csv("data2016.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2017 <- read_csv("data2017.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2018 <- read_csv("data2018.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))

data2019 <- read_csv("data2019.csv", col_types = cols(
  .default = "f",
  Time = col_skip(),
  Minutes = col_integer(),
  Seconds = col_integer(),
  Away = col_character(),
  Away_Points_Added = col_integer(),
  Away_Score = col_integer(),
  Home_Score = col_integer(),
  Diff_Home = col_integer(),
  Diff_Away = col_integer(),
  Diff_Abs = col_integer(),
  Home_Points_Added = col_integer(),
  Home = col_character()
))


data <- bind_rows(data2014,data2015, data2016, data2017, data2018, data2019)

Fourth_Plus <- c("4","OT1","OT2","OT3","OT4")

data <- data %>%
  filter(Quarter %in% Fourth_Plus) 

```

Add variable showing logical if play was home or away:
```{r}
data$Is_Home <- if_else(data$Play == data$Home_Court,"Home", "Away")
```


```{r}
data$Play <- ''

#combine away and home play details into one column
for(i in 1:nrow(data)) { 
  if (is.na(data$Away[i])) {
    data$Play[i] <- data$Home[i]
  } else { 
    data$Play[i] <- data$Away[i]
  } 
}

#log shot results per play
data$Shot_Result <- NA
data[which(str_detect(data$Play, ".*misses.*")), "Shot_Result"] <- 0
data[which(str_detect(data$Play, ".*makes.*")), "Shot_Result"] <- 1


#Extract Player

data$Player <- str_extract(data$Play, "\\b\\p{Lu}\\.\\s*\\w+")
data$Shot_Length <- str_extract(data$Play, "..\\sft.*")
data$Shot_Length <- str_extract(data$Shot_Length, ".\\d")


#is dunk? might help to illinate from clutch SHOOTING as opposed to clutch playing

data$Is_Dunk <- str_extract(data$Play, " 2-pt dunk")
data$Is_Dunk <- if_else(is.na(data$Is_Dunk), "No", "Yes")

#is three?
data$Is_Three <- str_extract(data$Play, " 3-pt jump shot")
data$Is_Three <- if_else(is.na(data$Is_Three), "No", "Yes")

#is free throw
data$Is_FT <- str_extract(data$Play, " free throw")
data$Is_FT <- if_else(is.na(data$Is_FT), "No", "Yes")

#score differential
data$Away_Score <- as.numeric(data$Away_Score)
data$Home_Score <- as.numeric(data$Home_Score)

for(i in 1:nrow(data)) { 
  if (is.na(data$Away_Points_Added[i])) {
    data$Points_Added[i] <- data$Home_Points_Added[i]
  } else { 
    data$Points_Added[i] <- data$Away_Points_Added[i]
  } 
}

#margin at time of shot
for(i in 1:nrow(data)) { 
  if (data$Is_Home[i] == "Home") {
    data$Lead[i] <- data$Home_Score[i] - data$Away_Score[i]
  } else { 
    data$Lead[i] <- data$Away_Score[i] - data$Home_Score[i]
  } 
}

data$Points_Added[which(is.na(data$Points_Added))] <- 0
```

# Filter Last 6 Years Playoff Data

```{r}

clutch_analysis <- data %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4, Quarter %in% Fourth_Plus) %>%
  filter(Is_FT == "No") %>%
  select(Year, Round, Game, Player, Home_Score, Away_Score, Diff_Abs, Lead, Quarter, Minutes,Seconds,Shot_Result,Points_Added,Shot_Length,Is_Dunk,Is_Three,Is_FT, Is_Home, Home_Court, On_Road, Play)

clutch_analysis$Shot_Length[which(is.na(clutch_analysis$Shot_Length),)] <- 0
       
```

# Assign non bums (stars) and current and past star closers

```{r}
stars <- c("S. Curry", "D. Lillard", "L. James","J. Harden", "K. Durant","K. Leonard", "R. Westbrook", "D. Wade","D. Nowitzki","D. Rose","T. Parker","T. Duncan", "K. Bryant", "A. Davis", "D. DeRozan","D. Mitchell","G. Antetokounmpo", "J. Butler", "J. Embid", "K. Garnett", "K. Irving", "K. Lowry","K. Thompson","K. Towns","K. Walker","N. Jokic","P. George","P. Pierce","V. Oladipo", "M. Ginobli", "B. Griffin", "B. Simmons", "C. McCollum")

stars_current <- c("S. Curry", "D. Lillard", "L. James","J. Harden", "K. Durant","K. Leonard","R. Westbrook") 

stars_yesterday <- c("D. Wade","D. Nowitzki","D. Rose","T. Parker","T. Duncan", "K. Bryant")

```

# Who Shoots Most Clutch Shots (Last 6 Years)

```{r}
clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 30) %>%
  group_by(Player) %>%
  summarise(sum =n())%>%
  arrange(desc(Player))


```


# Create Databases for All Specified Time Frames

```{r}
cl_30 <- clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 30)

cl_1 <- clutch_analysis %>%
  filter(Minutes == 0 & Seconds < 60)

cl_2 <- clutch_analysis %>%
  filter(Minutes < 2)

cl_3 <- clutch_analysis %>%
  filter(Minutes < 3)

cl_4 <- clutch_analysis %>%
  filter(Minutes < 4)

cl_5 <- clutch_analysis %>%
  filter(Minutes < 5)

#Figure out shots attempted and made by bums per time

cl_30 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_30 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_1 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_1 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_2 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_2 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_3 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_3 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_4 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_4 %>%
  filter(!Player %in% stars) %>%
  summarize(n())


cl_5 %>%
  filter(!Player %in% stars) %>%
  summarise(shots = sum(Shot_Result))
cl_5 %>%
  filter(!Player %in% stars) %>%
  summarize(n())

```

Results for bums below were used to manually insert into the Excel Doc "Clutch Complete" as a standard to compare the career percentages of the stars in that data set.

30: 67/156
1: 107/237
2: 163/394
3: 232/538
4: 281/703
5: 327/852


# Load Clutch Complete - Career Output of All Potential Star's Clutch Qualifying Shots

Clutch_Complete is a data set of a compilation of all supposed star players one would think is "clutch". It was compiled from NBA Advanced Stats and includes some 13 or so of the most likely "clutch" shooters today (and some from a while back like Kobe, Dirk, and Duncan, etc.). Again, this data includes each player's entire career output of clutch qualifying shots. The only exception is the pooled "Bums" entries which are only from the past 6 years of playoffs. These "bums" include anyone not a part of the 34 names in "stars" derived above. So not only is it not the superstars, but not the mid-level stars either. Just "bums" so to speak. I felt the past 6 years gives a good sample size while not going too far back to be potentially outdated. 

FYI: Notice Clutch_Complete turns into player_list below:

```{r message=FALSE}
library(readxl)
library(directlabels)
Clutch_Complete <- read_excel("Clutch Complete.xlsx")

player_list <- Clutch_Complete %>%
  select(Player, Min, FGM,FGA) %>%
  filter(!is.na(FGA)) %>%
  group_by(Player, Min) %>%
  summarize(tot_made = sum(FGM), tot_att = sum(FGA)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

#initial graph hiding the identity of the bums
ggplot(player_list[player_list$Player %in% stars_current | player_list$Player == "Bums",], aes(Min, Percentage, color = Player))+
  scale_x_continuous(limits = c(-.4,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink1","black", "red2", "gold2", "brown4", "purple4", "blue", "yellow4"))+
  geom_dl(aes(label = Player), method = list(dl.combine("first.points"), cex = 0.8))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

#Reveal the almighty bums
g <- ggplot(player_list[player_list$Player %in% stars_current | player_list$Player == "Bums",], aes(Min, Percentage, color = Player))+
  scale_x_continuous(limits = c(-.40,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink2","black", "red2", "gold2", "firebrick4", "purple4", "blue", "yellow4"))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

direct.label(g, "first.qp")

p <- ggplot(player_list[!player_list$Player %in% stars_current,], aes(Min, Percentage, color = Player))+
  geom_smooth(method = "loess")+
    scale_x_continuous(limits = c(0,5), breaks = c(0,.5,1,2,3,4,5))+
  scale_y_continuous(breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  geom_smooth(method = "loess", size = 2) +
  scale_color_manual(values = c("deeppink1","blue", "red2", "black", "purple4", "gray", "gray3"))+
  theme_classic()+
  theme(legend.position="none")+
  ggtitle("FG% When Tied or Team Down 3 or Less in Playoffs \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

direct.label(p, "first.qp")

```

# Star Low Sample size Issue

Table Showing Low Sample Size For Individual Stars vs. Many of Bums

```{r}
player_list %>%
  filter(Min == .5) %>%
  group_by(Player) %>%
  summarize(total = sum(tot_att))%>%
  arrange(desc(total))
  
```


Create a pooled star group so comparison with bums is possible:

```{r}

player_list$Type <- if_else(player_list$Player %in% stars,"Stars", "Bums")

player_list %>%
  filter(Min == .5) %>%
  group_by(Type) %>%
  mutate(Percent = sum(tot_made) / sum(tot_att)) %>%
  summarise(Avr = mean(Percent))

#USE IN ARTICLE

bum_star_graph <- player_list %>%
  filter(Player != "Stars") %>%
  group_by(Type, Min) %>%
  summarize(tot_made = sum(tot_made), tot_att = sum(tot_att)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

ggplot(bum_star_graph, aes(Min, Percentage, color = Type))+
  geom_smooth(se = FALSE, size = 2)+
  scale_y_continuous(limits = c(0,1), breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  scale_x_continuous(breaks = c(.5,1,2,3,4,5))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))
```

Same things as above only this time let's only focus on the BEST closers, leaving out the mid-level stars:

```{r}
super_stars <- c("S. Curry", "L. James", "K. Durant","K. Leonard", "D. Nowitzki","T. Duncan", "K. Bryant")

bum_star_graph <- player_list %>%
  filter(Player != "super_stars") %>%
  group_by(Type, Min) %>%
  summarize(tot_made = sum(tot_made), tot_att = sum(tot_att)) %>%
  mutate(Percentage = round(tot_made/tot_att,2)) %>%
  arrange(desc(Percentage))

ggplot(bum_star_graph, aes(Min, Percentage, color = Type))+
  geom_smooth(se = FALSE, size = 2)+
  scale_y_continuous(limits = c(0,1), breaks = c(.1,.2,.3,.4,.5,.6,.7,.8,.9))+
  scale_x_continuous(breaks = c(.5,1,2,3,4,5))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Super 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Across Time Remaining")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))
```



Great plot. clearly shows that as minutes wane, clutch shooters (ACROSS THEIR CAREER) choke more than bums (IN LAST 6 YEARS) but with 5 min stars are better as expected. 

# Last Minutes Comparison (Stars vs. Bums)
Now go into last minute - show each and every second. 

```{r}
clutch_analysis$Star_or_Bum <- NA
clutch_analysis$Star_or_Bum <- if_else(clutch_analysis$Player %in% stars,"Stars", "Bums")


last_min <- clutch_analysis %>%
  filter(Minutes == 0) %>%
  group_by(Star_or_Bum, Seconds) %>%
  summarize(tot_made = sum(Shot_Result), tot_att = n()) %>%
  mutate(Percentage = round(tot_made/tot_att,2))

ggplot(last_min, aes(Seconds, Percentage, fill = Star_or_Bum, color = Star_or_Bum))+
  geom_smooth(method = "loess", se = FALSE, size = 2)+
  scale_x_continuous(breaks = c(0,1,5,30,60))+
  scale_y_continuous(limits = c(0,1))+
  scale_color_manual(values = c("deeppink2","springgreen4"))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Final Minute")+
  ylab("Shooting Percentage")+
  xlab("Seconds Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

```



#Comparision of Final 12 Minutes by Each Individual Minute (Stars vs. Bums)

```{r}
time_perc_stars <- clutch_analysis %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4) %>%
  filter(Quarter == "4" | 
           Quarter == "OT1" | Quarter == "OT2" | 
           Quarter == "OT3" | Quarter == "OT4") %>%
  filter(Is_FT == "No" & Player %in% stars) %>%
  group_by(Minutes) %>%
  summarise(n = n(), Percentage = mean(Shot_Result))

time_perc_bums <- clutch_analysis %>%
  filter(Shot_Result == 0 | Shot_Result == 1) %>%
  filter(Lead < 1 & Lead > -4) %>%
  filter(Quarter == "4" | 
           Quarter == "OT1" | Quarter == "OT2" | 
           Quarter == "OT3" | Quarter == "OT4") %>%
  filter(Is_FT == "No" & !Player %in% stars) %>%
  group_by(Minutes) %>%
  summarise(n = n(), Percentage = mean(Shot_Result))

ggplot(time_perc_stars, aes(Minutes, Percentage))+
  geom_bar(stat = "identity", fill = "blue")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))

ggplot(time_perc_bums, aes(Minutes, Percentage))+
  geom_bar(stat = "identity", fill = "deeppink2")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))

#BEATUFUL PLOT!!! JUST MESS WITH COLORS TO FIND BEST MATCH. 
ggplot(NULL, aes(Minutes, Percentage))+
  geom_bar(aes(fill = "Percentage"), stat = "identity", data = time_perc_stars, alpha = .4, fill = "springgreen4")+
  geom_bar(aes(fill = "Percentage"), stat = "identity", data = time_perc_bums, alpha = .4, fill = "deeppink2")+
  scale_x_continuous(breaks = c(0,1,2,3,4,5,6,7,8,9,10,11))+
  theme_classic()+
  ggtitle("Pooled 'Closers' vs. Pooled Bums FG% \n Playoffs When Tied or Down 3 or Less \n Final 12 Minutes")+
  ylab("Shooting Percentage")+
  xlab("Minutes Left in Game")+
  theme(plot.title = element_text(hjust = 0.5))

```

Despite a dip in the 7 minute and 9 minute range, stars out-shot bums until there were less than 3 minutes to go overall. 

#Why Are Bums Better?

Since the crossover occurs under 3 minutes in %, use this as the parameter. All 34 stars used in this graph. You can see the disparity in 0-3 ft. shots in favor of the bums. Easier shots, better chance of making it. Not super difficult. 
```{r}

clutch_analysis$Shot_Length <- as.numeric(clutch_analysis$Shot_Length)


clutch_analysis %>%
  filter(Minutes < 3) %>%
  ggplot(aes(Shot_Length, fill = factor(Star_or_Bum)))+
  geom_bar(position = "dodge", alpha = .7)+
  scale_fill_manual("Stars Vs. Bums", values = c("Bums" = "deeppink2", "Stars" = "springgreen4"))+
  theme_classic()+
  ggtitle("Amount of Shots Per Foot \n Pooled 'Closers' vs. Pooled Bums in Playoffs When Tied or Down 3 or Less \n Final 3 Minutes")+
  scale_x_continuous(breaks = seq(0,62,2))+
  ylab("Number of Shots")+
  xlab("Length of Shots (in feet)")+
  theme(plot.title = element_text(hjust = 0.5))
```


